---
title: "GPS Analytics"
author: "GPS Analytics Group"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "Guided step by step analysis of income based mobility patterns from GPS data."
---

# Introduction

In latest years new sources of data have been proven to be valuable assets to understand human behavior and thus
help in the creation of public policies based on evidence. Among those lies mobile phone data which is commonly used
to study spatio-temporal patterns in the form of mobility or social connections. This type of data usually comes in two
flavors; Call Detail Record (CDR), which is basicaly the information generated by a telecom transaction between at least two
devices, i.e. phone calls or SMS and it can be used to determine connections between users and a location proxy given the
connection of the device to the nearest antenna. The second type is GPS data. In present times mobile phones, particularely smart phones,
which global penetration is estimated to be almost 50\%, are used for much more than Pier to Pier (P2P) transactions as they
are capable of connecting to the internet through apps. There are companies that partner up with app creators in order to
store information about the usage of their products, this way when used in a phone they store a token with a unique device ID
and time stamp plus any extra information that the user has consent to share, such as geolocation which is obtained via the
high precision GPS anthenas the phones are equiped with plus WiFi triangulation in some cases to upscale it.

We are interested in the latter source of information as the aim of our study is to untangle the universal disparities, if any,
in mobility reduction during the COVID-19 pandemic between socioeconomic groups. We focused our study in 6 middle-income countries (Brazil, Colombia,
Indonesia, Mexico, Philippines and South Africa).

A summary plot of our results is displayed in the interactive chart below. The first panel represents the percentage
change of the total time spent at home by socioeconomic group,
the second one is the percentage change of daily users commuting to work also by socioeconomic group and the last one
shows the percentage change of work commutes in the low-income group by the wealth of the neighborhood where the work is located.

```{r echo=FALSE}
knitr::include_url("chart.html", height = "400px")
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

A more detailed explanation of these results and the procedure to obtained them is described in the following chapters.
